---
title: "data_viz"
author: "Charles Hendrickson"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)

# specifically for data-viz
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)
library(sf)

```

## Get lizards data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

# Histogram of lizard weights 
```{r}
ggplot(data = lizards, aes(x = weight)) + 
  geom_histogram(fill = "orange",
                 color = "purple",
                 size = 0.2)
```


# total_length on x-axis
#weight on y axis


```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22,
             fill = "yellow",
             color = "red",
             size = 4,
             alpha = 0.5)
```

# Bubble plot - changing the size of a point based on the variable 
color of points based on common_name and size vares with total length 
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, size = total_length),
             shape = 22,
             ) +
  theme_light()

```

# Facet_wrap lines things up
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name, ncol = 7)

```


# facet_grid matches the intersections 
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail)
```


```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

# Same thing
lizard_counts <- lizards %>% dplyr::count(common_name)

lizard_count_cn_tail <- lizards %>% dplyr::count(common_name, tail)

```

```{r}
ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, n), x = n)) +
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(x = "lizard counts")
  
```

```{r}
lizard_counts_new <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, n))
```

## Axis scale breaks, limits and labels 

Scatterplot: total_length (x) versus weight (y)
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 10, 50, 500),
                     limits = c(0, 500),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0,0))
  
```
# Transform the date column to class Date, then find counts of observations by date.
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)
  
class(lizard_counts$date)

```

make a line plot (geom_line()) of date (x) and count (y)

```{r}
ggplot(data = lizard_counts, aes(x= date, y = n)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%y")
  
  
```











